---
title: "HW 5"
author: "Rifa Momin"
date: "2024-02-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(mosaic)
library(ggplot2)
library(knitr)
library(kableExtra)
letter_frequencies <- read.csv("letter_frequencies.csv")
```

# Problem 1

```{r, echo=FALSE, warning=FALSE, message=FALSE}

sim_ironbank= do(100000)*nflip(n=2021, prob=.024)
p_value_bank = (sum(sim_ironbank >= 70)) /100000


```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(sim_ironbank) + geom_histogram(aes(x = nflip), binwidth = 1, fill = "pink2", color = "black") +
  labs(title = "Distribution of Simulation", x = "Number of Flips", y = "Frequency")

```


The null hypothesis is that securities trades from Iron Bank are flagged at the same 2.4% baseline rate as that of other traders.

The test statistic is the number of flagged trades. In this data,there were 70 flagged trades out of 2021.

The calculate probability distribution, 100, 000 Monte Carlo simulations were ran, with each simulation having 2021 flips and a .024 probability.

After running the simulation, `r sum(sim_ironbank >= 70)` simulations yielded 70 cases of flagged trades or more. `r sum(sim_ironbank >= 70)` out 100,000 cases gives us a p-value of `r (sum(sim_ironbank >= 70)) / 100000`. Since `r (sum(sim_ironbank >= 70)) / 100000` < .05 suggests that the null hypothesis is rejected; Iron Bank has a greater baseline rate of flagged trades than 2.4%.

# Problem 2


```{r, echo=FALSE, warning=FALSE, message=FALSE}

sim_health= do(100000)*nflip(n=50, prob=.03)
sum(sim_health >= 8)
```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

ggplot(sim_health) + geom_histogram(aes(x = nflip), binwidth = 1, fill = "pink2", color = "black") +
  labs(title = "Distribution of Simulation", x = "Number of Flips", y = "Frequency")
```

The null hypothesis is that health code violations from Gourmet Bites are cited at the same 3% baseline rate of other restaurants in the city, on average.

The test statistic is the number of health code violations. In this data, there were 8 violations out of 50 inspections.

The calculate probability distribution, 100,000 Monte Carlo simulations were ran, with each simulation having 50 flips and a .03 probability.

After running the simulation, `r sum(sim_health >= 8)` simulations yielded 8 health code violations or more. `r sum(sim_health >= 8)` out 100,000 cases gives us a p-value of `r (sum(sim_health >= 8)) / 100000`. Since `r (sum(sim_health >= 8)) / 100000` < .05 suggests that the null hypothesis is rejected; Gourmet Bites  has a greater baseline rate of health code violations than the average restaurant in the city.




# Problem 3

## Part A


```{r, echo=FALSE, warning=FALSE, message=FALSE}
file_path <- "C:/Users/rifar/OneDrive/Desktop/UT 23-24/SDS 315/brown_sentences.txt"

brown_sentences <- readLines(file_path)
brown_sentences <- tibble(Sentence = brown_sentences)

```

```{r, echo=FALSE, warning=FALSE, message=FALSE}

calculate_chi_squared = function(brown_sentences, freq_table) {
  
  # remove all non-letter characters
  brown_sentences = gsub("[^A-Za-z]", "", brown_sentences) 
  brown_sentences = toupper(brown_sentences)
  
  observed_counts <- table(factor(strsplit(brown_sentences, "")[[1]], levels = freq_table$Letter))
  total_letters = sum(observed_counts)
  
  expected_counts = total_letters * freq_table$Probability
  
  chi_squared_stat = sum((observed_counts - expected_counts)^2 / expected_counts)
  
  return(chi_squared_stat)

}

chi_squared_stats <- list()

# loop through all sentences to compute chi-squared, store stat in empty list above
for (x in 1:nrow(brown_sentences)){
  chi_squared_stats[x] = calculate_chi_squared(brown_sentences[x,], letter_frequencies)
}

reference_distribution = unlist(chi_squared_stats)


ggplot(data = NULL) + geom_histogram(aes(x=unlist(reference_distribution)), binwidth = 3, color = "black", fill = "lightblue2") + labs(title = "Distribution of Chi Squared", x = "", y = "Frequency")
```


## Part B

```{r, echo=FALSE, warning=FALSE, message=FALSE}

# Define the sentences
sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)

```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
calculate_p_value <- function(observed_statistic, reference_distribution) {
  p_value <- sum(reference_distribution >= observed_statistic) / length(reference_distribution)
  return(p_value)
}
# store stats while looping through each sentence 
chi_squared_stats <- numeric(length = 10)
p_values <- numeric(length = 10)

for (i in 1:10) {
  chi_squared_stats[i] <- calculate_chi_squared(sentences[i], letter_frequencies)
  p_values[i] <- calculate_p_value(chi_squared_stats[i], reference_distribution)
}

p_values <- round(p_values, 3)

p_value_table <- data.frame(SentenceNum = 1:length(p_values), P_Value = p_values)
kable((p_value_table), booktabs = TRUE)

```


The watermarked sentence is sentence 6: Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland. The table above shows that it is the only sentence with a p-value (.009) less than .05, suggesting statistical significance when comparing the chi squared distribution of the sentence to the reference chi squared distribution.